---
title: 'Practise: ggplot2, dplyr, t test, Linear Regression'
output:
  html_document:
    df_print: paged
---

###### In this notebook we will investigate through a dataset named gapminder
###### You need to install gapminder, ggplot2, dplyr
###### First include the library gapminder and let's see the summary
###### we see that it has columns like country, continent, year, life expentency, Total population and GDP/capita


```{r}
library(gapminder)
data("gapminder")
summary(gapminder)
```
###### After calling this attach fucntion, we dont need to mention dataset again and again, suppose, instead of gapminder$column we can use only column

###### We will look the histogram of population, its looked messy
###### we go for a log transformation. whats a log transformation, when to appy, we might discuss later
###### we also look at the density of the graph it looks like a normal distribution

```{r}
attach(gapminder)
hist(pop)
hist(log(pop))
d <- density(log(pop))
plot(d)
```

###### Let's see boxplot for different continent life expectancy, WE see africa has the lowest life expectancy and Ocenia and Europe has the heighest
###### But in europe we have lots of outliers/exceptions

```{r}
boxplot(lifeExp~ continent)
```
###### Lets see gdp vs life expectency scatter plot, Its looks not perfect
###### Let go for log transformation, Its looks better?

###### And we also see a positive relation between gdp and life expectancy, More money, more lifetime?

```{r}
plot(lifeExp ~ gdpPercap)
plot(lifeExp ~ log(gdpPercap))
```

###### Let's see this numerically, correlation between lifeExp ~ log(gdpPercap))

```{r}
cor.test(lifeExp, log(gdpPercap))
```

###### We see that correlation is .80, thats a strong correlation. 
###### We will go for T test, to see if that things significant, And after that a linear regression
###### BUT, before that lets do some dplyr :) 
###### dplyr used pipe %>% (shortcut, ctrl+shift+m) function, pipe will attach previous condition/functionality to the next line


```{r}
library(dplyr)
gapminder %>%
  select(country, lifeExp)%>%
  # Select used for selection of only this columns e:g counry, lifeEXp
  filter(country == "Ireland" | country == "Bangladesh") %>% 
  # Then we filter out our country of interst
  group_by(country) %>% 
  # We group by them by country so that we can see the result countrywise
  summarise(avg_life = (mean(lifeExp)))
  #summarise function to seee the summary output

```

###### Now we will save this things to new dataframe for later investigation/visualization

```{r}
#library(dplyr)
df1<- gapminder %>%
  select(country, lifeExp)%>%
  # Select used for selection of only this columns e:g counry, lifeEXp
  filter(country == "Ireland" | country == "Bangladesh") %>% 
  # Then we filter out our country of interst
  group_by(country)

```


###### Lets go for a t test to see if the difference of the mean significant
###### Our null hypothesis is there is no change, Alternative is there is a change for mean value of these two country

```{r}
t.test(data = df1, lifeExp~country)
```
###### Look at the p value, its too low
###### If t is low(less than alpha which is .05), the null must go, There is no place for null hyp. here, so Alternative wins
###### There is a difference between two countrys mean
###### Look at the output again, we are 95 percent confident that the change is between 29.18 to 17.17




#### Let's have some plotting

```{r}
library(ggplot2)
```




###### Lots of things here, 
###### first ggplot, here data = gapminder, aesthetics, for x axis it is gdp/capita
###### for y axis its lifeExpectancy
###### Here col = year, that is design our plot with different color for different year
###### shape = continent, that is different shape for different continent
###### size = population, that larger bubbule/point for larger population
###### alpha = 0.7, that is transparancy, chage it between 0 to 1 to see the difference
###### geom_smooth, for a smooth line following the slope for different continent, method = lm, to make
###### the line a straight line, remove method = lm, now see the line is not straight anymore
###### and facet is used for separating a single graph to different graph
###### that is, facet_wrap, continent, make 5 graph for different continet
###### Remove facet part, now see all the graph joined together in a single graph


```{r}
ggplot(gapminder, aes(log(gdpPercap),lifeExp, col = year, shape=continent,  size = pop))+
  geom_point(alpha= 0.7)+geom_smooth(method = lm)+facet_wrap(~continent)

```

###### Most wanted Linear regression
###### We waited for it, as strong positive correlation, and change are significant

```{r}
lm(lifeExp~gdpPercap)
```

###### Linear regression, Y = mx +b (here, y = estimeted_output, m = slope, x = our input, b = coefficient/constant )
###### here m = 7.649e-04,  and b = 5.396e+01

```{r}
summary(lm(lifeExp~gdpPercap))
```
###### here Multiple R-squared is 0.34, that is not a good model mayb
###### It shows, our model can explain 34 percent of data only. 


##### Lets go for a log trasformation
```{r}
summary(lm(lifeExp~log(gdpPercap)))
```
###### Looks better, right? R squared increased to 65 percent :) 